This project was funded by the Department of Energy's Nuclear Energy University
Program. I am grateful for the financial support that allowed me to conduct my
research. In addition to our project sponsors, I have many others 
to thank for their support over the years. 

I owe Dr. Paul Wilson a immense amount of gratitude for his years of continued
academic support and guidance. You are the embodiment of what it means to be
a researcher, and educator, and a citizen. Your dedication to the well-being of your
students is something we should all aspire to. I can say with absolute
certainty that I would not be in the position I am today without your help. 
Thank you for advising me as a student in the Computational Nuclear Engineering 
Research Group (CNERG). My experience working with you has been nothing short of phenomenal.

I would like to thank Dr. Greg Nellis for his support of my work. 
Your advice was incredibly helpful as I worked out the kinks 
in my thermal hydraulic model. I also owe a thanks to Becky Sondelski, 
the graduate student responsible for the power cycle model development. 
I am thankful to have had the opportunity to work with you. Thank you to Mark
Anderson for your material science and corrosion expertise. Thank you Jake
Blanchard for agreeing to sit on my defense committee.

I would also like to thank my colleagues in CNERG. Thank you Young-Hui Park 
for taking me in as an undergraduate researcher assisting on your project. Thank you
Pat Shriwise and Baptiste Mouginot for your friendship and countless hours of
banter and laughter, it made the hard days a little less so. Thank you Katie
Mummah and Katie Biegel, I will forever cherish our camaraderie as we navigated
our first year of grad school together. Thank you Chelsea D'Angelo and Nancy
Granda-Duarte for your long conversations and commiserations in 434, I will be
forever grateful for our friendships. Every day on the fourth floor of ERB 
was a new adventure-both personally and professionally, to everyone in CNERG,
your consistent positive attitudes and support made working here a privilege. I
will always view our academic and professional relationships as overshadowed by
our personal friendships. I will miss this community dearly.

To the students, faculty, and staff, both past and present of the Engineering
Physics Department, thank you for making this a safe and enjoyable place to
learn and work. I have spent nearly 6 years as a member of this community and I
will cherish every moment for the rest of my life. The friendships and
professional connections I have made in my time here will continue to fulfill me
as I move on with my career. We have a truly special program and
an even better community here.

To my friend Sam Rebenstorf. Your friendship has been a relief from 
my academic frustrations over my entire time at UW. Thank you for being there
for me and supporting me as I worked towards my Master's degree.

To my sisters Libby and Mary. Thank you for your consistent cheerleading and
love over the years. Your company has always been a respite from the struggles
and monotony of everyday life. I learn so much from you both and can't express
how happy I am to be your brother.

To my parents Andy and Patty, thank you for your support, love, and guidance for
the past 24 years. You raised me to think for myself, to always stay curious,
and to never give up. I will always appreciate the sacrifices you made for 
Libby, Mary, and me.

Finally I would like to thank my girlfriend, Julia. Your companionship meant
the world to me as I fumbled my way through grad school. You have been
a steady source of support and comfort in my life and I could not be luckier
than I am with you. You provide me with constant motivation to push 
boundaries and become the best possible version of myself. 
I cannot wait to see what we accomplish together.
