\chapter{Conclusion and Future Work}

\section{Conclusion}
The goal of this project was to develop a constrained reactor mass model that
could be rapidly executed and integrated into a total power system mass
optimization scheme. The model needed to incorporate physics
constraints (cooling, criticality) in order to approximate and optimize reactor
mass. The model was designed to be integrated
into an overall power cycle model as part of a mass optimization process. A
framework was developed to model reactor mass as a submodule of an overall
power system mass model. The modeling framework considered both thermal
hydraulic and neutronic requirements to optimize a minimum-mass reactor for a
fixed set of flow inputs. The reactor mass submodel was successfully integrated
into the power cycle model and was used to optimize the overall mass of the
power cycle. The resultant reactor was verified using traditional reactor
physics tools to be a valid design that, with significant time, funding, and
engineering effort, could be developed into a full-fledged reactor concept.

\section{Future Work}
While the project was successful in setting up a workflow for modeling a nuclear
space reactor, there is much more work to be done. The workflow should be
applied to different designs to search for better mass performance and more
modeling should be done to improve the technical readiness of the designs.
Specifically, reactivity control and safety analysis needs to be incorporated
into the design process.

\subsection{Model Verification}
Some verification work was done to check the reactivity of an optimized reactor
design from the reactor mass model, more work should be done to verify the thermal
model. The thermal model made many simplifying assumptions such as: plane wall
conduction, 1D heat transfer, and a constant radius of conduction. Detailed
thermal analysis with Finite Element software would be useful to verify the
coolability of the optimized reactor designs.

\subsection{Expanded Mass Modeling}
Some reactor components were not considered for mass modeling. Notably absent
from the model were control elements and shielding. Shielding was not considered
important due to the deployment of the reactor on Mars. Martian regolith could
be used to shield the reactor. The reactor could be deployed in other parts of
the solar system or as a power system on a spaecraft. Shielding could be
important depending on the application. The mass of reactivity control systems
should be added to the reactor mass model. The mass of control systems should 
be relatively independent of power but should be added to the reactor mass
model.

\subsection{Expanded Design Exploration}
In order to constrain the reactor optimization space, two fuel types were
chosen. While this limitation was useful for developing the model, 
there are many other fuel types that could be considered. For monolithic
cores, this exploration is simple as developing new critical radius curves and
applying different thermal and physical properties. Switching fuel form would be
more complicated. Modified 1D heat transfer modeling would be required to
account for a different fuel geometry. It would be interesting to explore
TRISO fuel as well as high temperature graphite fuel forms. These fuels are
capable of operating at extremely high temperatures, potentially increasing
efficiency of the power cycle.

The reactor mass model and the supporting workflow was designed to support this
work. The geometry-specific calculations are modular and isolated in the thermal
hydraulic scripts. The core heat transfer functions could be modified to support
a new geometry. The neutronics modeling workflow was completely scripted in
Python, easily facilitating the replication of the modeling with a new geometry.

\subsection{Modeling Work}
Reactor designs from this modeling workflow yield very rough models and limited
information about a design. The designs were meant to provide estimates of mass
to help an overall power cycle mass optimization. Reactivity control and safety
analysis needs to be performed on any chosen design in order to advance the
design any further. Reactivity control will be essential to ensure the reactor
remains subcritical until reaching the mission destination. Once the power
system has been deployed, reactivity control systems will be essential for
reactor startup and any unexpected shutdowns during the system lifetime.
